Article 1: Digital Trace Data as Indicators of Social Data Validating Google Trends for  use in Scientific Research

### 2022-03-01 paper1.Rmd
Finish initial findings

---
title: "Article 1:	Construct Validity and Correspondence of Google Trends"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
output:
  bookdown::pdf_document2: 
    latex_engine: xelatex
    number_sections: true
  # bookdown::word_document2: 
  #   toc: true
  #   reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{longtable}
   - \usepackage{caption}
   - \usepackage{booktabs}
   - \usepackage{placeins}
   - \doublespacing
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
---

@@

```
```{r data}
vacc_hes <- read_rds(here('data', 'cleaned_data', 'vacc_hes.rds'))
mask <- read_rds(here('data', 'cleaned_data', 'mask.rds')) %>%  ungroup()
covid <- read_rds(here('data', 'cleaned_data', 'covid.rds'))
suicide <- read_rds(here('data', 'cleaned_data', 'suicide.rds'))
pres_2016 <- read_rds(here('data', 'cleaned_data', 'pres_2016.rds'))
pres_2020 <- read_rds(here('data', 'cleaned_data', 'pres_2020.rds'))
covid_corrs <- read_rds(here('data', 'from_hpc', 'covid_corrs.rds'))
suicide_corrs <- read_rds(here('data', 'from_hpc', 'suicide_corrs.rds'))

@@

@@

New big data sources have led to vast possibilities for social science
research because they are bigger, cheaper, and already available
[@kingEnsuringDataRichFuture2011; @lazerComputationalSocialScience2009;
@salganikBitBitSocial2017]. Before overenthusiastically embracing these
sources into our workflows, social scientists must clearly establish
parameters under which these data sources could be operationalized
[@bailCulturalEnvironmentMeasuring2014; @lazerParableGoogleFlu2014]. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" \[@lazerParableGoogleFlu2014,
p. 1203\]. Building on this prior research outlining these various
issues with big data [@boydCriticalQuestionsBig2012;
@lazerIssuesConstructValidity2015], this paper tests the construct
validity of Google Search Trends as an indicator of three different
cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators
from the NORC General Social Survey, United States county-level suicide
rates from National Center for Health Statistics, US cases rates of
Covid-19 from The New York Times, historical US Presidential Election
results, and the American National Election Survey. Data will be
analyzed against corresponding operationalized Google Trends
longitudinal data using Pearson's r for pairwise correlations, testing
for the strength of relationships between the Google Trend indicator and
the respective comparison indicator. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

"Behaviors and Attitudes","Vaccine Hesitancy for COVID-19", "Cross-Sectional", "March 3 – 15, 2021", "covid conspiracy', 'COVID-19 vaccine', 'Coronavirus', 'Covid-19'",

"Behaviors and Attitudes","Mask-Wearing Survey Data", "Cross-Sectional", "July 2 - 14, 2020", "'Face Mask', 'Mask', 'Cloth Face Mask'",

@@

"Health","County Suicide Rates", "Longitudinal", "Yearly 2010-2020", "'Suicide', 'Depression',' Suicide Hotline'",

@@

'Political',"Presidential Election Results", "Cross Sectional", "2016 & 2020", "'Hilary Clinton', 'Donald Trump', 'Joe Biden'"
 ) %>% as_grouped_data(groups = c("type"), columns = NULL)

@@

## Measures
### Google Trends
This paper focuses on a validation of Google search trends [@googletrends].
Google trends portray the search frequency for specific search terms across
designated media markets areas (DMAs), a nonoverlapping aggregation of U.S.
counties to 210 media markets based on similar population clusters [@dma_key].
Raw data is on a scale from 0 to 100, with 100 being the maximum search
popularity out of all DMAs. Google Search Trend are now only available
cross-sectionally (a single time period across a geography) or as time-series (a
single geo-location across time). To remedy this and build a longitudinal
dataset of each search topic for the longitudinal datasets, I follow the method
proposed in @park_etal (p. 5). This method involves building a dataset of
unscaled cross-sectional values, selecting a DMA to use to establish the
rescaling ratio (I use 'Los Angeles CA'), and then finding the time-series
values for the one DMA. To find the rescaling ratio for each week in the
time-series, you divide the time-series value for each week by the
cross-sectional value for each week, resulting in a rescaling vector to be used
for all weeks in the dataset across geographies. To rescale each longitudinal
value, multiply the respective week's rescaling ratio by the cross-sectional
value. Rescaled longitudinal data was compared against time-series data for
multiple test counties and was equivalent. For a more in-depth explanation of
this procedure, see @park_etal (p. 5). Missing datapoints in longitudinal
datasets were filled in with interpolated values using `zoo::na.approx()`
[@zoo].

@@

## Cultural
```{r corr-results}
#|  results='asis'

@@

table2 <- tribble(
  ~'grp', ~'variable', ~'measure',~'trend1', ~'trend2', ~'trend3', ~'trend4', 
  "Vaccine Hesitancy", NA,  NA, 'covid_conspiracy', 'covid_19_vaccine', 'coronavirus', 'covid_19', 
  "Vaccine Hesitancy", "Vaccine Hesitancy", "Pearson's R Correlation",
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$covid_conspiracy, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$covid_19_vaccine, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$coronavirus, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$covid_19, use = "complete.obs"),4)),

  "Mask Attitudes",NA,  NA, 'face_mask', 'mask', 'cloth_face_mask',  NA,
  "Mask Attitudes","Mask Rare", "Pearson's R Correlation",
  as.character(round(cor(mask$mask_rare, mask$face_mask, use = "complete.obs"),4)),
  as.character(round(cor(mask$mask_rare, mask$mask, use = "complete.obs"),4)),
  as.character(round(cor(mask$mask_rare, mask$cloth_face_mask, use = "complete.obs"), 4)), NA,

  "Covid Rates",NA,  NA, 'covid_19',  'smell_loss', 'taste_loss', NA, 
  "Covid Rates","Covid Rate", "Pearson's R Correlation", 
  as.character(round(cor(covid$covid_rate, covid$covid_19, use = "complete.obs"),4)),
  as.character(round(cor(covid$covid_rate, covid$smell_loss, use = "complete.obs"),4)),
  as.character(round(cor(covid$covid_rate, covid$taste_loss, use = "complete.obs"),4)),
  NA,
  "Covid Rates","Covid Rate", "rmcorr", 
  as.character(covid_corrs[[2,2]]), 
  as.character(covid_corrs[[2,3]]), 
  as.character(covid_corrs[[2,4]]), NA,

  "Suicide Rates",NA,  NA, 'suicide', 'depression', 'suicide_hotline', NA, 
  "Suicide Rates","Suicide Rate", "Pearson's R Correlation",
  as.character(round(cor(suicide$death_rate, suicide$suicide, use = "complete.obs"),4)),
  as.character(round(cor(suicide$death_rate, suicide$depression, use = "complete.obs"),4)),
  as.character(round(cor(suicide$death_rate, suicide$suicide_hotline, use = "complete.obs"),4)),
  NA,
  "Suicide Rates","Suicide Rate", "rmcorr", 
  as.character(suicide_corrs[[2,2]]), 
  as.character(suicide_corrs[[2,3]]), 
  as.character(suicide_corrs[[2,4]]), NA,

  "2016 Presidential Votes",NA,  NA, 'Hilary Clinton',  'Donald Trump', NA, NA,
  "2016 Presidential Votes","2016 Votes for Clinton", "Pearson's R Correlation",
  as.character(round(cor(pres_2016$clinton_p, pres_2016$`Hilary Clinton`, use = "complete.obs"),4)),
  as.character(round(cor(pres_2016$clinton_p, pres_2016$`Donald Trump`, use = "complete.obs"),4)), NA, NA,
  "2016 Presidential Votes","2016 Votes for Trump", "Pearson's R Correlation",
  as.character(round(cor(pres_2016$trump_p, pres_2016$`Hilary Clinton`, use = "complete.obs"),4)),
  as.character(round(cor(pres_2016$trump_p, pres_2016$`Donald Trump`, use = "complete.obs"),4)), NA, NA,

  "2016 Presidential Votes",NA,  NA, 'Joe Biden',  'Donald Trump', NA, NA,
  "2016 Presidential Votes","2020 Votes for Biden", "Pearson's R Correlation",
  as.character(round(cor(pres_2020$biden_p, pres_2020$`Joe Biden`, use = "complete.obs"),4)),
  as.character(round(cor(pres_2020$biden_p, pres_2020$`Donald Trump`, use = "complete.obs"),4)), NA, NA,
  "2016 Presidential Votes","2020 Votes for Trump", "Pearson's R Correlation",
  as.character(round(cor(pres_2020$trump_p, pres_2020$`Joe Biden`, use = "complete.obs"),4)),
  as.character(round(cor(pres_2020$trump_p, pres_2020$`Donald Trump`, use = "complete.obs"),4)), NA, NA
)  %>%   
  select(grp, measure, variable, trend1, trend2, trend3, trend4) %>% 
  as_grouped_data(groups = c("grp"), columns = NULL)

@@

The first attitudinal indicators is Vaccine Hesitancy for COVID-19 vaccines
[@vaches_data]. When comparing the measure of vaccine hesitancy to four
different Google Search Trends, the correlation does not exceed -0.4021
('Coronavirus' and vaccine hesitancy). Correlation's under |0.40| are considered
to be weak according to the common rules of thumb. When running these
correlations in multiple linear regression (see the results in table 
\@ref(tab:vacc_hes_analysis)), I see an $r^2$ of 0.221 (Model 1)
and 0.346 (Model 2), indicating that the Google Trends are able to explain about
22% of the variation in Vaccine Hesitancy alone. Demographic characteristics
like the percentage of households with broadband internet and the population
density are able to explain about 35% of the variation (model 3), outperforming
the first two models. The Trend coefficients themselves, however, are
significant and remain significant when controlling for demographics. This
reinforces the finding from the Pearson Correlation that there is a significant
but weak relationship between the Google Trends and Vaccine Hesitancy.

@@

vacc_hes <- vacc_hes %>% 
  select(-c(DMA_google, Geo_QName)) %>% 
  mutate(across(covid_conspiracy:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

vacc_hes_1  <- lm(vacc_hesitant  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband,
                  data = vacc_hes)
vacc_hes_2 <- lm(vacc_hesitant  ~  covid_19_vaccine +     
                   coronavirus +covid_19,
                 data = vacc_hes)
vacc_hes_2b <- lm(vacc_hesitant  ~ covid_conspiracy + covid_19_vaccine +
                    coronavirus +covid_19,
                  data = vacc_hes)
vacc_hes_3  <- lm(vacc_hesitant  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband + 
                    covid_19_vaccine +
                    coronavirus +covid_19,
                   data = vacc_hes)

modelsummary(list(vacc_hes_2, vacc_hes_2b, vacc_hes_1, vacc_hes_3), 
             # title = "Linear Regression Results for Vaccine Hesitancy",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE'
             ) %>%
  kable_styling(latex_options = c("hold_position", 'condensed'),
                font_size = 8)

@@

The second attitudinal measure I test is how Google Trends relates to rare mask
usage. As with vaccine hesitancy, the Pearson correlations are weak, if not 
negligible. I introduce these trends in multiple linear regression in table 
\@ref(tab:mask_analysis). Model 1 demonstrates that these three Google Search
Trends can explain about 9% of the variance in mask usage across U.S. counties,
reinforcing the conclusion that the relationship is quite weak. The coefficients
themselves are somewhat significant in Model 1. However, after including the
demographic varaibles in Models 2 and 3, we see that the relationship between
the Google Search Trends and mask usage is strengthed in magnitude and in
significance, likely indicating a repression effect due to background
relationships between the trends and demograhic variables. While the trends are
significant and match the magnitude of the demographic variables, the low $r^2$ 
for the trends still provides evidence that Google Search Trends data cannot 
replace survey analysis when trying to measure rare mask usage.

@@

mask <- mask %>% 
  select(-c(DMA_google, Geo_QName)) %>% 
  mutate(across(face_mask:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

mask_1  <- lm(mask_rare  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband,
                  data = mask)
mask_2 <- lm(mask_rare  ~  face_mask + mask + cloth_face_mask,
                 data = mask)
mask_3  <- lm(mask_rare  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband + 
                    face_mask + mask + cloth_face_mask,
                   data = mask)

@@

## Medical

```{r covid_analysis}
covid <- covid %>% 
  select(-c(dma_google, Geo_QName, county, state)) %>% 
  mutate(across(covid_19:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1])) %>% 
  left_join(covid %>%
              group_by(fips) %>%
              summarize(covid_rate_fips_mean = mean(covid_rate, na.rm = TRUE)),
            by = "fips") %>% 
  drop_na(covid_rate)

@@

modelsummary(list(covid_2, covid_1, covid_3), 
             # title = "Hierarchical Model for Covid Case Rates",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Random intercept per county'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE') %>%
  kable_styling(latex_options = c("hold_position", 'condensed'),
                font_size = 8)

@@

```
pearson correlations
rmcorr
rsquared
coefficients
conclusion

```{r suicide_analysis}
suicide <- suicide %>% 
  select(-c(dma_google, Geo_QName, county, population, deaths)) %>% 
  mutate(across(suicide:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1])) %>% 
  left_join(suicide %>%
              group_by(fips) %>%
              summarize(death_rate_fips_mean = mean(death_rate , na.rm = TRUE)),
            by = "fips") %>% 
  drop_na(death_rate )

suicide_1  <- lme(death_rate   ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income + broadband +
                  death_rate_fips_mean + year,
                random = ~ 1 | fips, na.action=na.omit,data = suicide)
suicide_2 <- lme(death_rate   ~  
                 suicide + depression + suicide_hotline + death_rate_fips_mean  + 
                 year,
               random = ~ 1 | fips, na.action=na.omit, data = suicide)
suicide_3  <- lme(death_rate   ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income + broadband + 
                  suicide + depression + suicide_hotline + death_rate_fips_mean  +
                  year,
                random = ~ 1 | fips, na.action=na.omit, data = suicide)

modelsummary(list(suicide_2, suicide_1, suicide_3), 
             # title = "Hierarchical Model for Suicide Rates",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Random intercept per county'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE') %>%
  kable_styling(latex_options = c("hold_position", 'condensed'),
                font_size = 8)

# TODO Drop SD intercept and observations,
# TODO why is the r2 so high? death_rate_fips_mean?
```
pearson correlations
rmcorr
rsquared
coefficients
conclusion

## Political

>>>>

---
title: "Article 1: Digital Trace Data as Indicators of Social Data Validating Google Trends for  use in Scientific Research"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
output:
  bookdown::pdf_document2: 
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
  # bookdown::word_document2: 
  #   toc: true
  #   reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{longtable}
   - \usepackage{caption}
   - \usepackage{booktabs}
   - \usepackage{placeins}
   - \usepackage{hhline}
   - \doublespacing
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
---

@@

```
```{r data}
#| cache = FALSE

vacc_hes <- read_rds(here('data', 'cleaned_data', 'vacc_hes.rds'))
mask <- read_rds(here('data', 'cleaned_data', 'mask.rds')) %>%  ungroup()
covid <- read_rds(here('data', 'cleaned_data', 'covid.rds'))
suicide <- read_rds(here('data', 'cleaned_data', 'suicide.rds'))
pres_2016 <- read_rds(here('data', 'cleaned_data', 'pres_2016.rds'))
pres_2020 <- read_rds(here('data', 'cleaned_data', 'pres_2020.rds'))
covid_corrs <- read_rds(here('data', 'from_hpc', 'covid_corrs.rds'))
suicide_corrs <- read_rds(here('data', 'from_hpc', 'suicide_corrs.rds'))

@@

Digital Trace Data as Indicators of Social Data: Validating Google Trends for
 use in Scientific Research

@@

New big data sources have led to vast possibilities for social science
research because they are bigger, cheaper, and already available
[@kingEnsuringDataRichFuture2011; @lazerComputationalSocialScience2009;
@salganikBitBitSocial2017]. Before overenthusiastically embracing these
sources into our workflows, social scientists must clearly establish
parameters under which these data sources could be operationalized
[@bailCulturalEnvironmentMeasuring2014; @lazerParableGoogleFlu2014]. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" [@lazerParableGoogleFlu2014,
p. 1203]. Building on this prior research outlining these various
issues with big data [@boydCriticalQuestionsBig2012;
@lazerIssuesConstructValidity2015], this paper tests the construct
validity of Google Search Trends as an indicator of three different
cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators
from the NORC General Social Survey, United States county-level suicide
rates from National Center for Health Statistics, US cases rates of
Covid-19 from The New York Times, historical US Presidential Election
results, and the American National Election Survey. Data will be
analyzed against corresponding operationalized Google Trends
longitudinal data using Pearson's r for pairwise correlations, testing
for the strength of relationships between the Google Trend indicator and
the respective comparison indicator. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

"Behaviors and Attitudes","Vaccine Hesitancy for COVID-19", "Cross-Sectional", "March 3 – 15, 2021", "Search Topics: 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Vaccine'",

"Behaviors and Attitudes","Mask-Wearing Survey Data", "Cross-Sectional", "July 2 - 14, 2020", "Search Topics: 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth Face Mask', 'Mask', 'Civil and Political Rights'",

@@

"Health","County Suicide Rates", "Longitudinal", "Yearly 2010-2020", "Search Topics: 'Suicide', 'Depression', Search Term:' Suicide Hotline'",

@@

'Political',"Presidential Election Results", "Cross Sectional", "2016 & 2020", "Search Topics: 'Hilary Clinton', 'Donald Trump', 'Joe Biden'"
 ) %>% as_grouped_data(groups = c("type"), columns = NULL)

@@

## Measures
### Google Trends
This paper focuses on a validation of Google search trends [@googletrends].
Google trends portray the search frequency for specific search terms across
designated media markets areas (DMAs), a nonoverlapping aggregation of U.S.
counties to 210 media markets based on similar population clusters [@dma_key].
Raw data is on a scale from 0 to 100, with 100 being the maximum search
popularity out of all DMAs. When available, I use Google search topics to
measure trends as opposed to search terms. Search topics are a more robust
measurement than a single search term: topics are aggregations of the rates of
multiple, highly correlated search terms together into a cohesive topic. For
example, while 'Beyoncé', 'Beyonce' and 'beyonce knowles' are all separate
search terms, 'Beyoncé Knowles' encompasses all of these into a single search
topic. Google Search Trend are only available cross-sectionally (a single time
period across a geography) or as time-series (a single geo-location across
time). To remedy this and build a longitudinal dataset of each search topic for
the longitudinal datasets, I follow the method proposed in @park_etal (p. 5).
This method involves building a dataset of unscaled cross-sectional values,
selecting a DMA to use to establish the rescaling ratio (I use 'Los Angeles
CA'), and then finding the time-series values for the one DMA. To find the
rescaling ratio for each week in the time-series, you divide the time-series
value for each week by the cross-sectional value for each week, resulting in a
rescaling vector to be used for all weeks in the dataset across geographies. To
rescale each longitudinal value, multiply the respective week's rescaling ratio
by the cross-sectional value. Rescaled longitudinal data was compared against
time-series data for multiple test counties and was equivalent. For a more
in-depth explanation of this procedure, see @park_etal (p. 5). Missing
datapoints in longitudinal datasets were filled in with interpolated values
using `zoo::na.approx()` [@zoo].

@@

## Validating Google Trends as metrics of attitudinal indicators

```{r corr-results}
#|  results='asis'

@@

table2 <- tribble(
  ~'grp', ~'variable', ~'measure',~'trend1', ~'trend2', ~'trend3', ~'trend4', ~'trend5', 
  "Vaccine Hesitancy", NA,  NA, 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Vaccine', NA,
  "Vaccine Hesitancy", "Vaccine Hesitancy", "Pearson's R Correlation",
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$covid_19_vaccine, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$coronavirus_disease, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$coronavirus_virus, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$vaccine, use = "complete.obs"),4)), NA,

  "Mask Attitudes",NA,  NA, 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth Face Mask', 'Mask', 'Civil and Political Rights',
  "Mask Attitudes","Mask Rare", "Pearson's R Correlation",
  as.character(round(cor(mask$mask_rare, mask$coronavirus_disease, use = "complete.obs"),4)),
  as.character(round(cor(mask$mask_rare, mask$coronavirus_virus, use = "complete.obs"),4)),
  as.character(round(cor(mask$mask_rare, mask$cloth_face_mask, use = "complete.obs"), 4)), 
  as.character(round(cor(mask$mask_rare, mask$mask, use = "complete.obs"), 4)), 
  as.character(round(cor(mask$mask_rare, mask$civil_and_political_rights, use = "complete.obs"), 4)),

  "Covid Rates",NA,  NA, 'covid_19',  'smell_loss', 'taste_loss', NA, NA, 
  "Covid Rates","Covid Rate", "Pearson's R Correlation", 
  as.character(round(cor(covid$covid_rate, covid$covid_19, use = "complete.obs"),4)),
  as.character(round(cor(covid$covid_rate, covid$smell_loss, use = "complete.obs"),4)),
  as.character(round(cor(covid$covid_rate, covid$taste_loss, use = "complete.obs"),4)),
  NA,NA, 
  "Covid Rates","Covid Rate", "repeated measures correlation coefficient", 
  as.character(covid_corrs[[2,2]]), 
  as.character(covid_corrs[[2,3]]), 
  as.character(covid_corrs[[2,4]]), NA,NA,

  "Suicide Rates",NA,  NA, 'suicide', 'depression', 'suicide_hotline', NA,NA,  
  "Suicide Rates","Suicide Rate", "Pearson's R Correlation",
  as.character(round(cor(suicide$death_rate, suicide$suicide, use = "complete.obs"),4)),
  as.character(round(cor(suicide$death_rate, suicide$depression, use = "complete.obs"),4)),
  as.character(round(cor(suicide$death_rate, suicide$suicide_hotline, use = "complete.obs"),4)),
  NA,NA, 
  "Suicide Rates","Suicide Rate", "repeated measures correlation coefficient", 
  as.character(suicide_corrs[[2,2]]), 
  as.character(suicide_corrs[[2,3]]), 
  as.character(suicide_corrs[[2,4]]), NA,NA,

  "2016 Presidential Votes",NA,  NA, 'Hilary Clinton',  'Donald Trump', NA, NA,NA, 
  "2016 Presidential Votes","2016 Votes for Clinton", "Pearson's R Correlation",
  as.character(round(cor(pres_2016$clinton_p, pres_2016$hillary_clinton_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2016$clinton_p, pres_2016$donald_trump_trend, use = "complete.obs"),4)), NA, NA,NA, 
  "2016 Presidential Votes","2016 Votes for Trump", "Pearson's R Correlation",
  as.character(round(cor(pres_2016$trump_p, pres_2016$hillary_clinton_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2016$trump_p, pres_2016$donald_trump_trend, use = "complete.obs"),4)), NA, NA,NA,

  "2020 Presidential Votes",NA,  NA, 'Joe Biden',  'Donald Trump', NA, NA,NA, 
  "2020 Presidential Votes","2020 Votes for Biden", "Pearson's R Correlation",
  as.character(round(cor(pres_2020$biden_p, pres_2020$joe_biden_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2020$biden_p, pres_2020$donald_trump_trend, use = "complete.obs"),4)), NA, NA,NA, 
  "2020 Presidential Votes","2020 Votes for Trump", "Pearson's R Correlation",
  as.character(round(cor(pres_2020$trump_p, pres_2020$joe_biden_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2020$trump_p, pres_2020$donald_trump_trend, use = "complete.obs"),4)), NA, NA, NA 
)  %>%   
  select(grp, measure, variable, trend1, trend2, trend3, trend4, trend5) %>% 
  as_grouped_data(groups = c("grp"), columns = NULL)

@@

The first attitudinal indicators is Vaccine Hesitancy for COVID-19 vaccines
[@vaches_data]. When comparing the measure of vaccine hesitancy to four
different Google Search Trends, the correlation does not exceed -0.4657
('Coronavirus (Disease)' and vaccine hesitancy). Correlation's under |0.40| are
considered to be weak according to the common rules of thumb. When running these
correlations in multiple linear regression (see the results in table
\@ref(tab:vacc_hes_analysis)), I see an $r^2$ of 0.233 (Model 1) and 0.411
(Model 2), indicating that the Google Trends are able to explain about 23% of
the variation in Vaccine Hesitancy alone. Demographic characteristics like the
percentage of households with broadband internet and the population density are
able to explain about 35% of the variation (model 2), outperforming the first
models. The Trend coefficients themselves, however, are significant and remain
significant when controlling for demographics. This reinforces the finding from
the Pearson Correlation that there is a significant but weak relationship
between the Google Trends and Vaccine Hesitancy.

@@

vacc_hes <- vacc_hes %>% 
  select(-c(DMA_google, Geo_QName)) %>% 
  mutate(across(covid_19_vaccine:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

vacc_hes_1  <- lm(vacc_hesitant  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband,
                  data = vacc_hes)
vacc_hes_2 <- lm(vacc_hesitant  ~  covid_19_vaccine + vaccine + 
                   coronavirus_disease + coronavirus_virus,
                 data = vacc_hes)
vacc_hes_3  <- lm(vacc_hesitant  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband + 
                    covid_19_vaccine + vaccine + 
                   coronavirus_disease + coronavirus_virus,
                   data = vacc_hes)

modelsummary(list(vacc_hes_2, vacc_hes_1, vacc_hes_3), 
             # title = "Linear Regression Results for Vaccine Hesitancy",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE'
             ) %>%
  kable_styling(latex_options = c("hold_position", 'condensed'),
                font_size = 8)

@@

The second attitudinal measure I test is how Google Trends relates to rare mask
usage. As with vaccine hesitancy, the Pearson correlations are negligible.
Correlation's under |0.20| are considered to be negigible according to the
common rules of thumb. I introduce these trends in multiple linear regression in
table \@ref(tab:mask_analysis). Model 1 demonstrates that these five Google
Search Trends can explain about 7% of the variance in mask usage across U.S.
counties, reinforcing the conclusion that the relationship is quite weak. The
coefficients themselves are somewhat significant in Model 1. However, after
including the demographic variables in Models 2 and 3, we see that the
relationship between the Google Search Trends and mask usage is strengthened in
magnitude and in significance, likely indicating a suppression effect due to
underlying relationships between the trends and demographic variables. While the
trends are significant and match the demographic variables in magnitude, the
low $r^2$ of 0.203 for the trends still provides evidence that Google Search
Trends data cannot replace survey analysis when trying to measure rare mask
usage.

@@

mask <- mask %>% 
  select(-c(DMA_google, Geo_QName)) %>% 
  mutate(across(coronavirus_disease:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

mask_1  <- lm(mask_rare  ~ 
                total_pop + pop_density + unemployment_rate + 
                over_65 + poverty_rate + median_income + broadband,
              data = mask)

mask_2 <- lm(mask_rare  ~ coronavirus_disease +
               coronavirus_virus + cloth_face_mask + 
               mask + civil_and_political_rights,
             data = mask)

mask_3  <- lm(mask_rare  ~ 
                total_pop + pop_density + unemployment_rate + 
                over_65 + poverty_rate + median_income + broadband + 
                coronavirus_disease +
                coronavirus_virus + cloth_face_mask + 
                mask + civil_and_political_rights,
              data = mask)

@@

## Validating Google Trends as metrics of health outcomes

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I test is
Covid-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of Covid-19. To further test the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of table \@ref(tab:covid_analysis) indicates that the trends themselves are able
to explain about 20% of the variation in Covid-19 rates ($r^2$ = 0.201). Few
of the demographic variables have an effect on Covid-19 rates in models 2 or 3,
though it seems there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to Covid-19 rates, but that this relationship
is weak and we should not attempt to use Google Trends as an indicator of actual
health outcomes.

```{r covid_analysis}
covid <- covid %>% 
  ungroup() %>% 
  select(-c(dma_google, Geo_QName, county, state)) %>% 
  mutate(across(covid_19:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1])) %>% 
  left_join(covid %>%
              group_by(fips) %>%
              summarize(covid_rate_fips_mean = mean(covid_rate, na.rm = TRUE)),
            by = "fips") %>% 
  drop_na(covid_rate)

@@

modelsummary(list(covid_2, covid_1, covid_3), 
             # title = "Hierarchical Model for Covid Case Rates",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Random intercept per county'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE|SD') %>%
  kable_styling(latex_options = c("hold_position", 'condensed'),
                font_size = 8)

@@

Another indicator of health outcomes I test to validate Google Trends as
possible health indicators are county suicide rates. As with previous cases,
there is only negligible correlation between actual suicide rates and Google
Search Trends, not exceeding 0.0845. In the more accurate repeated measures
correlation coefficient we see a maximum of 0.142 correlation, remaining
negligible. I introduce these trends into a random intercept model to control
for inter-group variation over time in table \@ref(tab:suicide_analysis). While
table \@ref(tab:suicide_analysis) model 1 reveals a significant relationship
between suicide hotline trends and suicide rates, the effect is small and the
entire model only has an $r^2$ of 0.058. Adding in demographic features in model
2 improves the $r^2$ quite a lot. In model 3, the inclusion of Google Trends
actually decreases the overall amount of suicide rate variance explained
compared to model 2. These analyses lead me to conclude that Google Search
Trends should not be used as indicators of suicide rates or intentions.

```{r suicide_analysis}
suicide <- suicide %>% 
  select(-c(DMA_google, Geo_QName, county, population, deaths)) %>%
  ungroup() %>% 
  mutate(across(death_rate:broadband, ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

suicide <- suicide %>% 
  left_join(suicide %>%
              group_by(fips) %>%
              summarize(death_rate_fips_mean = mean(death_rate , na.rm = TRUE)),
            by = "fips") %>% 
  drop_na(death_rate )

suicide_1  <- lme(death_rate   ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income +
                  # death_rate_fips_mean +
                    year,
                random = ~ 1 | fips, na.action=na.omit,data = suicide)
suicide_2 <- lme(death_rate   ~  
                 suicide + depression + suicide_hotline + 
                   # death_rate_fips_mean  +
                 year,
               random = ~ 1 | fips, na.action=na.omit, data = suicide)
suicide_3  <- lme(death_rate   ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income +
                  suicide + depression + suicide_hotline +
                    # death_rate_fips_mean  +
                  year,
                random = ~ 1 | fips, na.action=na.omit, data = suicide)
